Moving Towards Reproducible and Programmatic Generation of Neuroimaging Visualizations

### 2022-09-10 manuscript.Rmd
.

The visualization of neuroimaging data is one of the primary ways in which we evaluate data quality, interpret results, and communicate findings. These visualizations are commonly produced using graphical user interface-based (GUI) tools where individual images are opened and, within each instance, display settings are manually changed until the desired output is reached. In large part, the choice to use GUI-based software has been driven by a perception of convenience, flexibility, and accessibility. However, there now exist many code-based software packages which are well-documented and often do not require high-level knowledge of programming, making them more accessible to the neuroimaging community (Table 1). These tools are flexible and allow for the generation of reproducible, high-quality, and publication-ready brain visualizations in only a few lines of code (Figure 1), especially within the R, Python and MATLAB environments. The present commentary argues for the wide-spread adoption of code-generated visualizations by highlighting major advantages in replicability, flexibility, and integration over GUI based tools. We also providing didactic examples of visualizations and associated code as a point of reference.

@@

While many code-based tools are well documented and do not require a strong knowledge of programming, there can still be a steeper learning curve for new users, compared to using a GUI. This is especially true for the purpose of a publication-ready figure, were fine adjustments to visual auxiliary such as legend placement, font size and multi-panel figure positioning are needed. While most code-based tools offer some control over these finer steps, there are differences between them in feature availability and usability. Relatedly, while some interactive image viewers which can be opened within an integrated development environment like R-Studio (e.g. Muschelli, 2016), for quick and interactive viewing of single images, GUI tool can be faster and more practical.

>>>>

The visualization of neuroimaging data is one of the primary ways in which we evaluate data quality, interpret results, and communicate findings. These visualizations are commonly produced using graphical user interface-based (GUI) tools where individual images are opened and, within each instance, display settings are manually changed until the desired output is reached. In large part, the choice to use GUI-based software has been driven by a perception of convenience, flexibility, and accessibility. However, there now exist many code-based software packages which are well-documented and often do not require high-level knowledge of programming, making them more accessible to the neuroimaging community (Table 1). These tools are flexible and allow for the generation of reproducible, high-quality, and publication-ready brain visualizations in only a few lines of code (Figure 1), especially within the R, Python and MATLAB environments. Here we review the rational for the wide-spread adoption of code-generated visualizations by highlighting major advantages in replicability, flexibility, and integration over GUI based tools. We also providing didactic examples of visualizations and associated code as a point of reference.

@@

While many code-based tools are well documented and do not require a strong knowledge of programming, there can still be a steeper learning curve for new users, compared to using a GUI. This is especially true for the purpose of a publication-ready figure, were fine adjustments to visual auxiliary such as legend placement, font size and multi-panel figure positioning are needed. While most code-based tools offer some control over these finer steps, there are differences between them in feature availability and usability. Relatedly, while some interactive image viewers can be opened within an integrated development environment like R-Studio (e.g. Muschelli, 2016), for quick and interactive viewing of single images, GUI tool can be faster and more practical.
